Capstone Movielens Report

### 2025-11-17 reports/capstone-movielens.site/05.appendix.Rmd
Describing helper function (in progress): model.tune.param_range.

The `dvs` is a divisor used to calculate the step size for partitioning the interval into a sequence of $\lambda$ values, among which the value that minimizes *RMSE* has to be found.

So the step size of the sequence is calculated as follows:
$$
step = \frac{end - start}{dvs}
$$

The sequence obtained as a result of the transformation is equivalent to the one generated by the function [base::seq](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/seq) as follows:
```{r eval=FALSE}
seq(start, end, step)
```  
In fact, the `seq` function is called internally to generate the sequence during the execution of the algorithm.
:::

@@

::: { #func.model.tune.param_range.details .sidebar }
##### **Details** 
\
During execution, the algorithm performs the following steps:

  1. Creates a vector containing a sequence of $\lambda$ values generated as described in the [loop_starter](#func.model.tune.param_range.args.loop_starter) argument description above.

>>>>

The `dvs` is a divisor used to calculate the *step size* for partitioning the interval into a sequence of $\lambda$ values, among which the value that minimizes *RMSE* has to be found.

So the *step size* of the sequence is calculated as follows:
$$
step = \frac{end - start}{dvs}
$$

The sequence obtained as a result of the transformation is equivalent to the one generated by the function [base::seq](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/seq) as follows:
```{r eval=FALSE}
seq(start, end, step)
```

::: {.noteblock data-latex=""}
In the function body, the *step size* is stored in the local variable `seq_increment`, and the values of $\start$, $\end$, and $\dvs$ are stored in the local variables `seq_start`, `seq_end`, and `interval_divisor`, respectively.
:::

In fact, the `seq` function is called internally to generate the sequence during the algorithm's execution.

Thus, the sequence is generated by the following [line of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L256):
```{r eval=FALSE}
    test_param_vals <- seq(seq_start, seq_end, seq_increment)
```
:::

@@

::: { #func.model.tune.param_range.details .sidebar }
##### **Details** 
\
During execution, the algorithm organizes a cycle consisting of the following steps:

  1. Calculates the divisor to calculate the step size for partitioning the interval into a sequence of $\lambda$ values as described in the [loop_starter](#func.model.tune.param_range.args.loop_starter) argument description above using the following [line of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L240):
```{r eval=FALSE}
    seq_increment <- (seq_end - seq_start)/interval_divisor 
```

  2. Checks if the *step size* of the generated sequence is less than the minimum allowed value (in the current implementation, we use `1e-13`), and if `TRUE`, breaks out of the loop:
```{r eval=FALSE}
    if (seq_increment < 0.0000000000001) {
      warning("Function `model.tune.param_range`:
parameter value increment is too small.")

      put_log2("Function `model.tune.param_range`:
Final best RMSE for `parameter value = %1`: %2",
               param_values.best_result["param.best_value"],
               param_values.best_result["best_RMSE"])

      put(param_values.best_result)
      # browser()
      break
    }
```

::: {.noteblock data-latex=""}
In the code snippet above, we use the custom function [put_log2](#func.put_log2) described above in the [Logging Functions] section of *This Appendix*.
:::

  3. Creates a vector containing a sequence of $\lambda$ values generated as described in the [loop_starter](#func.model.tune.param_range.args.loop_starter) argument description above.
